Modelling the gas kinematics of an atypical Lyman alpha emitting compact dwarf galaxy

### 2017-03-28 paper/paper.tex
intro + mnras style

%Document Class
\documentclass[twocolappendix]{latex/emulateapj}

%Packages and Commands
\include{latex/packages_commands}

@@

\title{A simple model to describe how outflows and rotation shape the Lyman-$\alpha$
  emission line in galaxies}
\shorttitle{Rotation and outflows in the Lyman-$\alpha$ lines}

\shortauthors{Remolina-Gutierrez and Forero-Romero.}

\author{Maria Camila Remolina-Gutierrez, Jaime E. Forero-Romero} 
\affil{Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
  No. 18A-10, Edificio Ip, Bogot\'a, Colombia} 
\email{mc.remolina197@uniandes.edu.co}
\email{je.forero@uniandes.edu.co}

\keywords{Galaxies: high-redshift, Lyman Alpha Emission, Galaxy
  Rotation, Galaxy Outflows, Radiative Transfer}

\begin{abstract}
\noindent Young galaxies in the Universe have a strong \lya emission
caused by the ionized Hydrogen atoms in their interstellar
medium. When the spectrum of a galaxy has an intense peak around the
\lya natural frequency ($2.46\times 10^{15}$ Hz) it is called a Lyman
Alpha Emitter (LAE). Typical LAEs are very distant ($z \gtrsim
2$). This makes that all the data astronomers can obtain from them is
their spectra, and from there all the physical information of the
galaxy must be derived. Trying to solve this task requires the
creation of a simplified and solid model. In this paper we propose to
consider LAEs as a spherical distribution of Hydrogen atoms that
undergoes a solid body rotation and a radial expansion due to
outflows. We use radiative transfer Monte Carlo simulations to
interpret the Lyman-$\alpha$ line morphology. The main conclusion is
that this new model reproduces LAEs observed features in a clear way
and with consistent physical parameters. We also show results of
adjusting observational data for some selected objects to models with
and without bulk rotation. We finalize by discussing the possible
implications for these results in terms of the energetics required for
supernova feedback and outflows in high redshift galaxies. \\

REVISAR AL FINAL...\\
\end{abstract}

@@

@@

There are 3 parameters in this model that define a LAE: the rotational velocity (\vrot), the outflow velocity (\vout) and the optical depth (\tauh). \vout is due to material ejected from the galaxy, by supernovas(\cite{Verhamme06}, \cite{Orsi12}, \cite{Hashimoto2015}, \cite{Gronke2015}). \tauh roughly corresponds to number of H atoms found by a \lya photon if one traces a line from the center of the galaxy to its edge, and it resembles the mass of the LAE.\\

@@

These 3 parameters, \vrot, \vout and \tauh, leave the idea of a model of LAE, that although simple, considers the main galaxy's dynamics. All of them have been previously proposed and used by different authors, but never combined together (\cite{Adams72}, \cite{Harrington73}, \cite{Neufeld90}, \cite{Dijkstra06}, \cite{Verhamme06}, \cite{Forero12}, \cite{Martin2015}, \cite{Garavito14}, \cite{Neufeld91}, \cite{Laursen09}, \cite{Barnes11}, \cite{Verhamme12}, \cite{Yajima12}).\\

@@

%The process that takes place from the moment the first photon is emitted to the moment the last photon escapes the sphere, is called a simulation. Each simulation runs in the scientific computational cluster of Universidad de los Andes due to the computing resources it requires. Depending on \tauh, we need a different number of processors in order to minimize the time demands. For \tauh $=10^5, 10^6, 10^7$ we use 6, 12, 24 processors, respectively. The running times go from 7 hours to 1 week approximately, depending on the case. \\

@@

To summarize, the parameters influencing the spectra are \vrot, \vout, \tauh and $\theta$. In the next section we evaluate the impact each of these has on the resulting profile. \\

@@

\subsection{Resulting simulated spectra}
In order to define the ranges of \tauh, \vrot and \vout, it is necessary to refer to observational constraints. The common values for typical LAEs that were set as parameters are in Tab. \ref{tab:values}. We run CLARA's modified version for all the permutations of these 3 parameters. \\

@@

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=0.48\textwidth]{./figures/summary}
	\end{center}
	\caption{\textbf{4 different \lya profiles:} With \tauh$=10^5$ and $\theta \simeq 90^\circ$. The rotational velocity \vrot increases to the right and the outflows velocity \vout increases downwards. The intensity is in arbitrary units.
		\label{fig:summary}}
\end{figure}

@@

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=0.5\textwidth]{./figures/angles}
	\end{center}
	\caption{\textbf{\lya profile for different $\theta$.} For \tauh$=10^5$, \vrot$=50$ \kms and \vout$=20$ \kms. For \tauh$=10^6$, \vrot$=100$ \kms and \vout$=5$ \kms. For \tauh$=10^7$, \vrot$=100$ \kms and \vout$=15$ \kms. The intensity is in arbitrary units.
		\label{fig:angles}}
\end{figure}

@@

\begin{itemize}
	\item \tauh induces a redshift. Increasing the optical depth separates the line of the zero velocity line. \\
	\item \vout decreases the right peak's intensity. Higher \vout makes the left peak smaller until it merges with the right one. \\
	\item \vrot broadens the line and decreases the maximum intensity. Higher \vrot implies a flatter spectrum. This effect has not been deeply studied in literature. Only Garavito et al. \cite{Garavito14} has simulated its effect. Our results are consistent with their conclusions.
	\item $\theta$ increases the central valley of the line. Higher $\theta$ makes more emission at \lya natural frequency. \\ 
\end{itemize}

@@

The rotation effect induces a Doppler shift in frequency and displaces the only-outflow spectra. As seen in Fig. \ref{fig:rotation_doppler_outflow} rotation takes the resulting spectrum with \vrot$=0$ and displaces it from its central location. Then it weights the lines and they all merge into the resulting red \lya line.\\

@@

Finally, all the spectra produced from this new LAE model is roughly consistent with observations. The consideration of adding the new rotation parameter to the standard only-outflow model found in literature is reasonable and very powerful. \vrot and its consequent viewing angle $\theta$ can modify the morphology of the \lya line in new different ways by keeping the model simplified.\\

@@

	\item The effects of \vrot, \vout and \tauh are consistent with the different authors that have used them. \vrot broadens the \lya line. \vout increases the peaks asymmetry and \tauh induces a redshift around the zero velocity.

@@

\bibliographystyle{latex/apj}
\bibliography{references}

>>>>

\documentclass[a4,useAMS,usenatbib,usegraphicx]{mn2e} 
%=========================================================================
\usepackage{amsmath} 
\usepackage{amssymb} 
\usepackage {graphicx}
\usepackage[dvips]{epsfig}
\usepackage{epsfig}  
\usepackage{color}
\usepackage[normalem]{ulem}
\usepackage{hyperref}
\usepackage{caption}
%Non reposionated tables
\usepackage{float}
\restylefloat{table}
%Multiple columns support for tables
\usepackage{array}
\usepackage{booktabs}
\setlength{\heavyrulewidth}{1.5pt}
\setlength{\abovetopsep}{4pt}

%=========================================================================
%		INTERNAL MACROS
%=========================================================================
\def\be{\begin{equation}}
\def\ee{\end{equation}}
\def\ba{\begin{eqnarray}}
\def\ea{\end{eqnarray}}

% To highlight comments 
\definecolor{red}{rgb}{1,0.0,0.0}
\newcommand{\red}{\color{red}}
\definecolor{darkgreen}{rgb}{0.0,0.5,0.0}
\newcommand{\SRK}[1]{\textcolor{darkgreen}{\bf SRK: \textit{#1}}}
\newcommand{\SRKED}[1]{\textcolor{darkgreen}{\bf #1}}
\newcommand{\before}[1]{\textcolor{red}{ #1}}
\newcommand{\after}[1]{\textcolor{darkgreen}{ #1}}
\newcommand{\tol}{Tololo 1214-277}
\newcommand{\lya}{Ly$\alpha$}
\newcommand{\hb}{H$\beta$}
\newcommand{\ha}{H$\alpha$}
\newcommand{\oiii}{[OIII]}
\newcommand{\oii}{[OII]}
\newcommand{\nii}{[NII]}
\newcommand{\esca}{erg cm$^{-2}$ s$^{-1}$ \AA$^{-1}$}
\newcommand{\esc}{erg cm$^{-2}$ s$^{-1}$}
\newcommand{\es}{erg s$^{-1}$}
\newcommand{\esa}{erg s$^{-1}$}
\newcommand{\kms}{km s$^{-1}$}
\newcommand{\apj}{ApJ}  
\newcommand{\jcap}{JCAP}  
\newcommand{\apjs}{ApJS}  
\newcommand{\pasa}{PASA}  
\newcommand{\apjl}{ApJL}  
\newcommand{\aj}{AJ}  
\newcommand{\mnras}{MNRAS}  
\newcommand{\mnrassub}{MNRAS accepted}  
\newcommand{\aap}{A\&A}  
\newcommand{\aaps}{A\&AS}  
\newcommand{\araa}{ARA\&A}  
\newcommand{\nat}{Nature}  
\newcommand{\physrep}{PhR}
\newcommand{\pasp}{PASP}    
\newcommand{\pasj}{PASJ}    
\newcommand{\sigmaclump}{$54.3\pm 0.6$ km s$^{-1}$}
\newcommand{\inftyclump}{$54.3\pm 5.1$ km s$^{-1}$}
\newcommand{\probaclump}{$0.96\pm 0.01$}
\def\simgt{\lower.5ex\hbox{\gtsima}}
\def\simlt{\lower.5ex\hbox{\ltsima}}

@@

%=========================================================================
%		FRONT MATTER
%=========================================================================
\title[An atypical \lya\ dwarf galaxy]{
Modelling the gas kinematics of an atypical Lyman alpha emitting compact dwarf galaxy}
\author[J.E. Forero-Romero et al.]{
  Maria Camila Remolina-Guti\'errez$^1$ \&
  Jaime E. Forero-Romero $^{1}$ \thanks{je.forero@uniandes.edu.co}\\
  %%
  $^1$ Departamento de F\'isica, Universidad de los Andes, Cra. 1
  No. 18A-10 Edificio Ip, CP 111711, Bogot\'a, Colombia \\
}

       %\vspace*{6pt}\\

       %}
       \maketitle

\begin{abstract}
  Star-forming Compact Dwarf Galaxies (CDGs) resemble the expected
  pristine conditions of the first galaxies in the Universe and
  are the best systems to test models on primordial
galaxy formation and evolution.    
Here we report on one of such CDGs,

\end{abstract}

\begin{keywords}
galaxies: dwarf --- galaxies: individual:\tol\ --- radiative transfer --- Methods: numerical 
\end{keywords}

%=========================================================================
%		PAPER CONTENT
%=========================================================================

%*************************************************************************

@@

\bibliographystyle{mn2e}
\bibliography{references}

@@

There are 3 parameters in this model that define a LAE: the rotational velocity (vrot), the outflow velocity (vout) and the optical depth (tauh). vout is due to material ejected from the galaxy, by supernovas(\cite{Verhamme06}, \cite{Orsi12}, \cite{Hashimoto2015}, \cite{Gronke2015}). tauh roughly corresponds to number of H atoms found by a \lya photon if one traces a line from the center of the galaxy to its edge, and it resembles the mass of the LAE.\\

@@

These 3 parameters, vrot, vout and tauh, leave the idea of a model of LAE, that although simple, considers the main galaxy's dynamics. All of them have been previously proposed and used by different authors, but never combined together (\cite{Adams72}, \cite{Harrington73}, \cite{Neufeld90}, \cite{Dijkstra06}, \cite{Verhamme06}, \cite{Forero12}, \cite{Martin2015}, \cite{Garavito14}, \cite{Neufeld91}, \cite{Laursen09}, \cite{Barnes11}, \cite{Verhamme12}, \cite{Yajima12}).\\

@@

%The process that takes place from the moment the first photon is emitted to the moment the last photon escapes the sphere, is called a simulation. Each simulation runs in the scientific computational cluster of Universidad de los Andes due to the computing resources it requires. Depending on tauh, we need a different number of processors in order to minimize the time demands. For tauh $=10^5, 10^6, 10^7$ we use 6, 12, 24 processors, respectively. The running times go from 7 hours to 1 week approximately, depending on the case. \\

@@

To summarize, the parameters influencing the spectra are vrot, vout, tauh and $\theta$. In the next section we evaluate the impact each of these has on the resulting profile. \\

@@

\subsection{Resulting simulated spectra}
In order to define the ranges of tauh, vrot and vout, it is necessary to refer to observational constraints. The common values for typical LAEs that were set as parameters are in Tab. \ref{tab:values}. We run CLARA's modified version for all the permutations of these 3 parameters. \\

@@

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=0.48\textwidth]{./figures/summary}
	\end{center}
	\caption{\textbf{4 different \lya profiles:} With tauh$=10^5$ and $\theta \simeq 90^\circ$. The rotational velocity vrot increases to the right and the outflows velocity vout increases downwards. The intensity is in arbitrary units.
		\label{fig:summary}}
\end{figure}

@@

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=0.5\textwidth]{./figures/angles}
	\end{center}
	\caption{\textbf{\lya profile for different $\theta$.} For tauh$=10^5$, vrot$=50$ \kms and vout$=20$ \kms. For tauh$=10^6$, vrot$=100$ \kms and vout$=5$ \kms. For tauh$=10^7$, vrot$=100$ \kms and vout$=15$ \kms. The intensity is in arbitrary units.
		\label{fig:angles}}
\end{figure}

@@

\begin{itemize}
	\item tauh induces a redshift. Increasing the optical depth separates the line of the zero velocity line. \\
	\item vout decreases the right peak's intensity. Higher vout makes the left peak smaller until it merges with the right one. \\
	\item vrot broadens the line and decreases the maximum intensity. Higher vrot implies a flatter spectrum. This effect has not been deeply studied in literature. Only Garavito et al. \cite{Garavito14} has simulated its effect. Our results are consistent with their conclusions.
	\item $\theta$ increases the central valley of the line. Higher $\theta$ makes more emission at \lya natural frequency. \\ 
\end{itemize}

@@

The rotation effect induces a Doppler shift in frequency and displaces the only-outflow spectra. As seen in Fig. \ref{fig:rotation_doppler_outflow} rotation takes the resulting spectrum with vrot$=0$ and displaces it from its central location. Then it weights the lines and they all merge into the resulting red \lya line.\\

@@

Finally, all the spectra produced from this new LAE model is roughly consistent with observations. The consideration of adding the new rotation parameter to the standard only-outflow model found in literature is reasonable and very powerful. vrot and its consequent viewing angle $\theta$ can modify the morphology of the \lya line in new different ways by keeping the model simplified.\\

@@

	\item The effects of vrot, vout and tauh are consistent with the different authors that have used them. vrot broadens the \lya line. vout increases the peaks asymmetry and tauh induces a redshift around the zero velocity.

@@
